---
title: "Data analysis"
date: "12/09/2023"
output: html_document
---

## Libraries

```{r library, message = FALSE}
library(tidyverse)
library(readr)
library(boot)
library(table1)
library(gridExtra)
```


## Data Clean

```{r data load}
breastcancer_data =
  read_csv("Project_2_data.csv") |>
  janitor::clean_names()
  
summary(breastcancer_data)
```

```{r data clean}
bc = breastcancer_data |>
  mutate(
    race=case_when(
      race == "White" ~ 1,
      race == "Black" ~ 2,
      race == "Other" ~ 3),
    marital_status=case_when(
      marital_status == "Married" ~ 1,
      marital_status == "Divorced" ~ 2,
      marital_status == "Single" ~ 3,
      marital_status == "Widowed" ~ 4,
      marital_status == "Separated" ~ 5),
    t_stage=case_when(
      t_stage == "T1" ~ 1,
      t_stage == "T2" ~ 2,
      t_stage == "T3" ~ 3,
      t_stage == "T4" ~ 4),
    n_stage=case_when(
      n_stage == "N1" ~ 1,
      n_stage == "N2" ~ 2,
      n_stage == "N3" ~ 3),
    x6th_stage=case_when(
      x6th_stage == "IIA" ~ 1,
      x6th_stage == "IIIA" ~ 2,
      x6th_stage == "IIIC" ~ 3,
      x6th_stage == "IIB" ~ 4,
      x6th_stage == "IIIB" ~ 5),
    differentiate=case_when(
      differentiate == "Poorly differentiated" ~ 1,
      differentiate == "Moderately differentiated" ~ 2,
      differentiate == "Well differentiated" ~ 3,
      differentiate == "Undifferentiated" ~ 4),
    grade=case_when(
      grade == "1" ~ 1,
      grade == "2" ~ 2,
      grade == "3" ~ 3,
      grade == "anaplastic; Grade IV" ~ 4),
    a_stage=case_when(
      a_stage == "Regional" ~ 1,
      a_stage == "Distant" ~ 0),
    estrogen_status=case_when(
      estrogen_status == "Positive" ~ 1,
      estrogen_status == "Negative" ~ 0),
    progesterone_status=case_when(
      progesterone_status == "Positive" ~ 1,
      progesterone_status == "Negative" ~ 0),
    status=case_when(
      status == "Alive" ~ 1,
      status == "Dead" ~ 0)
    ) 
```

## Descriptive statistics for all variables

```{r Descriptive statistics}
summary(bc)
```

We change race, marital_status, t_stage, n_stage, x6th_stage, differentiate, and grade into multiple numeric levels, while a_stage, estrogen_status, progesterone_status, and status to binary levels. The above variables are categorical variables.

And age, tumor_size, regional_node_examined, reginol_node_positive, and survival_months are numeric variables.

## Covariance and Correlation

```{r correlation}
plot(bc)

cor(bc) |>
  knitr::kable(digits=4,caption="Correlation for all variables")
```

## Exploratory visualisation

```{r age}
plot1age = 
  breastcancer_data|>
  ggplot(aes(x = age)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.5), binwidth = 5) +
  theme_minimal() +
  labs(
    title = "Age Distribution",
    x = "Age",
    y = "Frequency"
  )
plot1age
```

```{r race}
plot2race = 
  breastcancer_data|>
  ggplot(aes(x = race)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Race Distribution",
    x = "Race",
    y = "Frequency"
  )
plot2race
```

```{r marital}
plot3marital = 
  breastcancer_data|>
  ggplot(aes(x = marital_status)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Marital Status Distribution",
    x = "Marital Status",
    y = "Frequency"
  )

plot3marital
```

```{r tstage}
plot4tstage = 
  breastcancer_data|>
  ggplot(aes(x = t_stage)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "T Stage Distribution",
    x = "T Stage",
    y = "Frequency"
  )

plot4tstage
```

```{r nstage}
plot5nstage = 
  breastcancer_data|>
  ggplot(aes(x = n_stage)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "N Stage Distribution",
    x = "N Stage",
    y = "Frequency"
  )

plot5nstage
```

```{r x6thstage}
plot6x6thstage = 
  breastcancer_data|>
  ggplot(aes(x = x6th_stage)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "x6th Stage Distribution",
    x = "x6th Stage",
    y = "Frequency"
  )

plot6x6thstage
```

```{r differentate}
plot7differentate = 
  breastcancer_data|>
  ggplot(aes(x = differentiate)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Differentate Distribution",
    x = "Differentate Group",
    y = "Frequency"
  )

plot7differentate
```

```{r grade}
plot8grade = 
  breastcancer_data|>
  ggplot(aes(x = grade)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Grade Distribution",
    x = "Grade",
    y = "Frequency"
  )

plot8grade
```

```{r astage}
plot9astage = 
  breastcancer_data|>
  ggplot(aes(x = a_stage)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "A_stage Distribution",
    x = "A Stage",
    y = "Frequency"
  )

plot9astage
```


```{r tumorsize}
plot10tumorsize = 
  breastcancer_data|>
  ggplot(aes(x = tumor_size)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Tumor Size Distribution",
    x = "Tumor Size",
    y = "Frequency"
  )

plot10tumorsize
```

```{r estrogen}
plot11estrogen = 
  breastcancer_data|>
  ggplot(aes(x = estrogen_status)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Estrogen Status Distribution",
    x = "Estrogen Status",
    y = "Frequency"
  )

plot11estrogen
```

```{r progesterone}
plot12progesterone = 
  breastcancer_data|>
  ggplot(aes(x = progesterone_status)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Progesterone Status Distribution",
    x = "Progesterone Status",
    y = "Frequency"
  )

plot12progesterone
```

```{r nodeexamined}
plot13nodeexamined = 
  breastcancer_data|>
  ggplot(aes(x = regional_node_examined)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Regional Node Examined Distribution",
    x = "Examined Regional Node",
    y = "Frequency"
  )

plot13nodeexamined
```

```{r nodepositive}
plot14nodepositive = 
  breastcancer_data|>
  ggplot(aes(x = reginol_node_positive)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Regional Node Positive Distribution",
    x = "Positive Reginol Node",
    y = "Frequency"
  )

plot14nodepositive
```

```{r survive months}
plot15survivalmonths = 
  breastcancer_data|>
  ggplot(aes(x = survival_months)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Survival Months",
    x = "Number of survival months",
    y = "Frequency"
  )

plot15survivalmonths
```

### Summarized plots

```{r numeric variables}
grid.arrange(plot1age, plot10tumorsize, plot13nodeexamined, 
             plot14nodepositive, plot15survivalmonths, ncol = 2)
```

We van see that age is approximately normal, while tumor size, regional node examined, regional node positive and survival months are skewed. We should do transforamtion on these four variables.

```{r categorical variables}
grid.arrange(plot2race, plot3marital,plot4tstage, 
             plot5nstage, plot6x6thstage,plot7differentate,
             plot8grade,plot9astage,plot11estrogen,plot12progesterone, ncol = 3)
```

## Transformation

```{r log transformation on numeric}
par(mfrow = c(2, 2)) 
hist(log(bc$tumor_size)) 
hist(log(bc$regional_node_examined)) 
hist(log(bc$reginol_node_positive))
hist(log(bc$survival_months))
```

We can see that tumor size and regional node examined become approximately normal after log transformation, while the other two still extremely skewed.

## Indicator Test

```{r Indicator Test}

```

## Model Fitting

### Initial Model

```{r linear fit}
lmfit=lm(status ~ ., data = bc) 
summary(lmfit)


par(mfrow=c(2,2))
plot(lmfit)
```

```{r logistic fit}
glmfit=glm(status ~ ., data = bc)
summary(glmfit)
par(mfrow=c(2,2))
plot(glmfit)
```

### Transformation model

```{r transformation on numeric}
newbc = bc |>
  mutate(ln_tumor=log(tumor_size),
         ln_examined=log(regional_node_examined)) |>
  dplyr::select(-tumor_size) |>
  dplyr::select(-regional_node_examined)
newbc
```

```{r new glm}
newglmfit=glm(status ~ ., data = newbc)
summary(newglmfit)
par(mfrow=c(2,2))
plot(glmfit)
```

3. transformation
3.1 interaction transformation 
3.2 partial test 
3.3 diagnostic boxcox 
4. Step-wise: forward/ backward /AIC 
5. final model 
6. model assumption (check multicolinearity (VIF))
7. cross validation
