---
title: "Data analysis"
date: "12/09/2023"
output: pdf_document
---

## Libraries

```{r library, message = FALSE}
library(tidyverse)
library(readr)
library(boot)
library(table1)
library(gridExtra)
library(MASS)
library(car)
library(dplyr)
library(leaps)
library(corrplot)
```

## Data Clean

```{r data load}
breastcancer_data =
  read_csv("Project_2_data.csv") |>
  janitor::clean_names()
  
summary(breastcancer_data)
```

```{r data clean}
bc = breastcancer_data |>
  mutate(
    race=case_when(
      race == "White" ~ 1,
      race == "Black" ~ 2,
      race == "Other" ~ 3),
    marital_status=case_when(
      marital_status == "Married" ~ 1,
      marital_status == "Divorced" ~ 2,
      marital_status == "Single" ~ 3,
      marital_status == "Widowed" ~ 4,
      marital_status == "Separated" ~ 5),
    t_stage=case_when(
      t_stage == "T1" ~ 1,
      t_stage == "T2" ~ 2,
      t_stage == "T3" ~ 3,
      t_stage == "T4" ~ 4),
    n_stage=case_when(
      n_stage == "N1" ~ 1,
      n_stage == "N2" ~ 2,
      n_stage == "N3" ~ 3),
    x6th_stage=case_when(
      x6th_stage == "IIA" ~ 1,
      x6th_stage == "IIIA" ~ 2,
      x6th_stage == "IIIC" ~ 3,
      x6th_stage == "IIB" ~ 4,
      x6th_stage == "IIIB" ~ 5),
    differentiate=case_when(
      differentiate == "Poorly differentiated" ~ 1,
      differentiate == "Moderately differentiated" ~ 2,
      differentiate == "Well differentiated" ~ 3,
      differentiate == "Undifferentiated" ~ 4),
    grade=case_when(
      grade == "1" ~ 1,
      grade == "2" ~ 2,
      grade == "3" ~ 3,
      grade == "anaplastic; Grade IV" ~ 4),
    a_stage=case_when(
      a_stage == "Regional" ~ 1,
      a_stage == "Distant" ~ 0),
    estrogen_status=case_when(
      estrogen_status == "Positive" ~ 1,
      estrogen_status == "Negative" ~ 0),
    progesterone_status=case_when(
      progesterone_status == "Positive" ~ 1,
      progesterone_status == "Negative" ~ 0),
    status=case_when(
      status == "Alive" ~ 1,
      status == "Dead" ~ 0)
    ) 
```

## Descriptive statistics for all variables

```{r Descriptive statistics}
summary(bc)
```
We change race, marital_status, t_stage, n_stage, x6th_stage, differentiate, and grade into multiple numeric levels, while a_stage, estrogen_status, progesterone_status, and status to binary levels. The above variables are categorical variables.

And age, tumor_size, regional_node_examined, reginol_node_positive, and survival_months are numeric variables.

## Covariance and Correlation

```{r correlation}
plot(bc)

cor(bc) |>
  knitr::kable(digits=4,caption="Correlation for all variables")
```

### Another plot for correlation

```{r plot for correlation}
corrplot(cor(bc), method = "square", type = "upper", diag = FALSE)

```

## Exploratory visualisation

```{r age}
plot1age = 
  breastcancer_data|>
  ggplot(aes(x = age)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.5), binwidth = 5) +
  theme_minimal() +
  labs(
    title = "Age Distribution",
    x = "Age",
    y = "Frequency"
  )
#plot1age
```

```{r race}
plot2race = 
  breastcancer_data|>
  ggplot(aes(x = race)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Race Distribution",
    x = "Race",
    y = "Frequency"
  )
#plot2race
```

```{r marital}
plot3marital = 
  breastcancer_data|>
  ggplot(aes(x = marital_status)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Marital Status Distribution",
    x = "Marital Status",
    y = "Frequency"
  )

#plot3marital
```


```{r tstage}
plot4tstage = 
  breastcancer_data|>
  ggplot(aes(x = t_stage)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "T Stage Distribution",
    x = "T Stage",
    y = "Frequency"
  )

#plot4tstage
```


```{r nstage}
plot5nstage = 
  breastcancer_data|>
  ggplot(aes(x = n_stage)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "N Stage Distribution",
    x = "N Stage",
    y = "Frequency"
  )

#plot5nstage
```

```{r x6thstage}
plot6x6thstage = 
  breastcancer_data|>
  ggplot(aes(x = x6th_stage)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "x6th Stage Distribution",
    x = "x6th Stage",
    y = "Frequency"
  )

#plot6x6thstage
```

```{r differentate}
plot7differentate = 
  breastcancer_data|>
  ggplot(aes(x = differentiate)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Differentate Distribution",
    x = "Differentate Group",
    y = "Frequency"
  )

#plot7differentate
```

```{r grade}
plot8grade = 
  breastcancer_data|>
  ggplot(aes(x = grade)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Grade Distribution",
    x = "Grade",
    y = "Frequency"
  )

#plot8grade
```

```{r astage}
plot9astage = 
  breastcancer_data|>
  ggplot(aes(x = a_stage)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "A_stage Distribution",
    x = "A Stage",
    y = "Frequency"
  )

#plot9astage
```


```{r tumorsize}
plot10tumorsize = 
  breastcancer_data|>
  ggplot(aes(x = tumor_size)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Tumor Size Distribution",
    x = "Tumor Size",
    y = "Frequency"
  )

#plot10tumorsize
```

```{r estrogen}
plot11estrogen = 
  breastcancer_data|>
  ggplot(aes(x = estrogen_status)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Estrogen Status Distribution",
    x = "Estrogen Status",
    y = "Frequency"
  )

#plot11estrogen
```

```{r progesterone}
plot12progesterone = 
  breastcancer_data|>
  ggplot(aes(x = progesterone_status)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Progesterone Status Distribution",
    x = "Progesterone Status",
    y = "Frequency"
  )

#plot12progesterone
```


```{r nodeexamined}
plot13nodeexamined = 
  breastcancer_data|>
  ggplot(aes(x = regional_node_examined)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Regional Node Examined Distribution",
    x = "Examined Regional Node",
    y = "Frequency"
  )

#plot13nodeexamined
```

```{r nodepositive}
plot14nodepositive = 
  breastcancer_data|>
  ggplot(aes(x = reginol_node_positive)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Regional Node Positive Distribution",
    x = "Positive Reginol Node",
    y = "Frequency"
  )

#plot14nodepositive
```

Y1 = survive months; (numeric)

Y2 = status; (binary)

```{r survive months}
plot15survivalmonths = 
  breastcancer_data|>
  ggplot(aes(x = survival_months)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "Survival Months",
    x = "Number of survival months",
    y = "Frequency"
  )

plot15survivalmonths
```

```{r status}
plot16status = 
  bc|>
  ggplot(aes(x = status)) +
  geom_bar(color = "blue", fill = alpha("blue", 0.5))+
  theme_minimal() +
  labs(
    title = "status",
    x = "status",
    y = "Frequency"
  )

plot16status
```

### Summarized plots for covariates

```{r numeric variables}
grid.arrange(plot1age, plot10tumorsize, plot13nodeexamined, 
             plot14nodepositive, ncol = 2)
```

We can see that age is approximately normal, while tumor size, regional node examined, and regional node positive are skewed. 

```{r categorical variables}
grid.arrange(plot2race, plot3marital,plot4tstage, 
             plot5nstage, plot6x6thstage,plot7differentate,
             plot8grade,plot9astage,plot11estrogen,plot12progesterone, ncol = 3)
```

## Transformation

We know that the tumor size, regional node examined, and regional node positive are skewed. We should do transformation on these variables. Before the transformation, we can use the Box-Cox plot to check which transformation work the best for them.

```{r boxcox}
bc_transform_tumorsize <- boxcox(breastcancer_data$tumor_size ~ 1, lambda = seq(-2, 2, by=0.1))
bc_transform_regionalnode_examined <- boxcox(breastcancer_data$regional_node_examined ~ 1, lambda = seq(-2, 2, by=0.1))
bc_transform_regionalnode_pos <- boxcox(breastcancer_data$reginol_node_positive ~ 1, lambda = seq(-2, 2, by=0.1))
```

The lambda value of tumor size is close to 0, so we should use log transformation, while the lambda value of regional node examined is around 0.5, we should take a square root to the value, and the lambda value of regional node positive is around -0.5, so we should take a take square root and take an (-1) exponent for transformation. 

```{r transformation on numeric}
par(mfrow = c(2, 2)) 
hist(log(bc$tumor_size)) 
hist(sqrt(bc$regional_node_examined)) 
hist(bc$reginol_node_positive**(-0.5))
```

We can see that tumor size and regional node examined become approximately normal after log transformation, while the regional node positive is still extremely skewed. Therefore, we may consider not using the variable of reginol_node_positive. 

### Transformation model

```{r}
newbc = bc |>
  mutate(ln_tumor=log(tumor_size),
         sqrt_examined=sqrt(regional_node_examined)) |>
  dplyr::select(-tumor_size) |>
  dplyr::select(-regional_node_examined) |>
  dplyr::select(-survival_months)
newbc
```
## chi-square tets
```{r}
# Create an empty data frame to store chi-square test results
chi_square_results <- data.frame(Variable = character(), 
                                 ChiSquare = numeric(), 
                                 PValue = numeric(), 
                                 stringsAsFactors = FALSE)

# Specify the dependent variable
dependent_var <- newbc$status

# List of independent variables
independent_vars <- newbc[,-13]

# Perform chi-square test for each independent variable
for (var in names(independent_vars)) {
  # Create a contingency table
  table <- table(dependent_var, newbc[[var]])
  
  # Perform chi-square test and suppress warnings for expected count less than 5
  test <- try(suppressWarnings(chisq.test(table)), silent = TRUE)
  
  # Store the results if the test was successful
  if (!inherits(test, "try-error")) {
    chi_square_results <- rbind(chi_square_results, data.frame(Variable = var, 
                                                               ChiSquare = test$statistic, 
                                                               PValue = test$p.value))
  } else {
    # Store NA if the test failed due to too many zero counts
    chi_square_results <- rbind(chi_square_results, data.frame(Variable = var, 
                                                               ChiSquare = NA, 
                                                               PValue = NA))
  }
}

chi_square_results
```
Based on the above chi-squared table, each variable listed has been tested for independence with respect to the dependent variable, and each shows a significant relationship.

## Indicator Test

### When y is status

```{r}
# indicator test when y is status
categorical_vars <- c("race", "marital_status", "t_stage", "n_stage", "x6th_stage", 
                       "differentiate", "grade", "a_stage", 
                       "estrogen_status", "progesterone_status")

newbc[categorical_vars] <- lapply(newbc[categorical_vars], factor)

formula <- as.formula("status ~ race + marital_status + t_stage + n_stage + x6th_stage + 
                       differentiate + grade + a_stage + estrogen_status + progesterone_status+ln_tumor + sqrt_examined+ reginol_node_positive+ age")

model <- glm(formula, data = newbc, family = binomial())

summary(model)
```
Based on the above indicator test summary, we delete grade and x6th_stage because their output was NA in the output linear model, since NA indicates these predicts may contribute collinearity. 

## Model Fitting

### Initial Model

```{r}
glmfit <- glm(status ~ race + marital_status + t_stage + n_stage + differentiate + 
               a_stage + estrogen_status + progesterone_status + ln_tumor + 
               sqrt_examined + reginol_node_positive + age, 
               data = newbc, family = binomial)
summary(glmfit)
par(mfrow = c(2, 2))
plot(glmfit)
```

## Indicator Variables categorize by Race

```{r race categorize}
# According to our data formating in the upper procedure, in Race: 
# 1 represents "white" 
# 2 represents "black"
# 3 represents "others"
## A newdataset with "white" as reference

# Ensure race is a factor and not an ordered factor
bc_ref <- newbc |> 
  mutate(race = factor(race, ordered = FALSE))

# Relevel to make white as reference
bc_ref <- bc_ref |> 
  mutate(race = relevel(race, ref = 1))

## Run a SLR for race only
single_race = glm(status ~ race, family = binomial, bc_ref)

summary(single_race)


## Run a MLR with race ref but without interaction terms
model_ref = glm(status ~ race + marital_status + t_stage + n_stage + differentiate + 
               a_stage + estrogen_status + progesterone_status + reginol_node_positive + ln_tumor + 
               sqrt_examined + age, 
               data = bc_ref, family = binomial)
#model_ref
summary(model_ref)
```

```{r}

## Run a logistic linear with race ref and with interaction terms
model_refinter = glm(status ~ race + marital_status + t_stage + n_stage + differentiate + 
               a_stage + estrogen_status + progesterone_status + reginol_node_positive + ln_tumor + 
               sqrt_examined + age + 
                race*(marital_status + t_stage + n_stage + differentiate + 
               a_stage + estrogen_status + progesterone_status + reginol_node_positive + ln_tumor + 
               sqrt_examined + age), 
               data = bc_ref, family = binomial)
#model_refinter
summary(model_refinter)


```

```{r}
## Run a MLR with race ref and with interaction terms selected
model_inter = glm(status ~ race + marital_status + t_stage + n_stage + differentiate + 
               a_stage + estrogen_status + progesterone_status + reginol_node_positive + ln_tumor + 
               sqrt_examined + age + 
                race*(marital_status + progesterone_status + ln_tumor),
               data = bc_ref, family = binomial)
#model_inter
summary(model_inter)
```
- From the SRL for single race variable, we can see that the race group of black has a P-value of 6.95*10^(-6) and other race group has a P-value of 0.0257 both smaller than 0.05, so there are statistically significant difference between that to the reference of group white. 
- From the MRL model without interaction terms, we can see that the AIC is around 2997.4 which is smaller than the original full model, when for both race2 and race3 we have a P-value smaller than 0.05 indicating significance, so we considering adding interaction terms.
- From the MRL model with race interaction terms added, we can see that the AIC is around 3016 which is getting larger, so we may have too many unnecessary interaction terms, and from the P-values we can see most the interaction covariates are insignificant except some terms in: race:marital_status, race:progesterone_status, race:ln_tumor.
- Therefore, we have the 3rd MRL model with race interaction terms selected, and we get a model with AIC of 2990.3 dropped sharply and the interaction term is included. 



## Step-wise: both direction/AiC


### Step-wise for MLR without interaction term:

```{r}
set.seed(123)
# Full Model
full_model <- glm(status ~ race + marital_status + t_stage + n_stage + differentiate + 
                  a_stage + estrogen_status + progesterone_status + reginol_node_positive + 
                  ln_tumor + sqrt_examined + age, 
                  data = newbc, family = binomial)

# Minimal Model (Intercept Only)
min_model <- glm(status ~ 1, data = newbc, family = binomial)

# Forward Selection
forward_model <- step(min_model, scope = list(lower = min_model, upper = full_model), 
                      direction = "forward", trace = FALSE)

# Backward Elimination
backward_model <- step(full_model, direction = "backward", trace = FALSE)

# Both Directions
stepwise_model <- step(min_model, scope = list(lower = min_model, upper = full_model), 
                       direction = "both", trace = FALSE)

# Print the summary of the chosen model
summary(forward_model)
# or summary(backward_model)
summary(backward_model)
# or summary(stepwise_model)
summary(stepwise_model)

```

From the both direction step-wise, the best model without interaction term has lowest AIC of 2993.5 is: status ~ n_stage + progesterone_status + differentiate + t_stage + age + race + estrogen_status + sqrt_examined + reginol_node_positive


### Step-wise for MLR with interaction term:

```{r}

# Full Model with Interaction Terms
full_interaction_model <- glm(status ~ race + marital_status + t_stage + n_stage + differentiate + 
               a_stage + estrogen_status + progesterone_status + reginol_node_positive + ln_tumor + 
               sqrt_examined + age + 
                race*(marital_status + progesterone_status + ln_tumor), 
                              data = bc_ref, family = binomial)

# Minimal Model (Intercept Only)
min_model <- glm(status ~ 1, data = bc_ref, family = binomial)

# Stepwise Selection (Forward, Backward, or Both)
stepwise_interaction_model <- step(min_model, scope = list(lower = min_model, upper = full_interaction_model), 
                                   direction = "both", trace = FALSE)

# Print the summary of the chosen model
summary(stepwise_interaction_model)
```

From the both direction step-wise, the best model with interaction term has lowest AIC of 2987.9 is: status ~ n_stage + progesterone_status + differentiate + t_stage + age + race + estrogen_status + sqrt_examined + reginol_node_positive + marital_status + progesterone_status:race + race:marital_status.




### VIF
```{r}
# Calculate VIF for the model without interaction
vif_result_bic <- vif(stepwise_model)
print(vif_result_bic)

# Calculate VIF for the interaction model
vif_result_aic <- vif(stepwise_interaction_model, type = 'predictor')
print(vif_result_aic)


```

## Partial Test 

### Partial Test for binary Y

```{r}
# Model without Interaction Terms
model_no_interaction <- glm(status ~ n_stage + progesterone_status + differentiate + t_stage + 
                            age + race + estrogen_status + sqrt_examined + reginol_node_positive, 
                            data = newbc, family = binomial)

# Model with Selected Interaction Terms
model_with_interaction <- glm(status ~ n_stage + progesterone_status + differentiate + t_stage + 
                              age + race + estrogen_status + sqrt_examined + reginol_node_positive + 
                              marital_status + progesterone_status:race + race:marital_status, 
                              data = bc_ref, family = binomial)

# Partial Test
anova(model_no_interaction, model_with_interaction, test = "Chisq")

```

The p-value is  below the common alpha level of 0.05, indicating that the inclusion of interaction terms in the model significantly improves the fit compared to the model without interaction terms. This result supports selecting the model with interaction terms, as it provides a better explanation of the variation in the response variable (status).


## ANOVA

#### anova for binary Y
```{r}
anova_result1 <- anova(model_with_interaction)
print(anova_result1)
summary(model_with_interaction)
```

```{r}
null_model <- glm(status ~ 1, data = bc_ref, family = binomial)

anova(null_model, model_with_interaction, test = "Chisq")
```

With p significant p value, we have enough evidence to show that the full model is better, which is status ~ n_stage + progesterone_status + differentiate + t_stage + age + race + estrogen_status + sqrt_examined + reginol_node_positive + marital_status + progesterone_status:race + race:marital_status

## Criterion procedures



### Model with Interaction terms
```{r}
aic_selected_bin_model <- stepAIC(model_with_interaction, direction = "both")
summary(aic_selected_bin_model)

bic_selected_bin_model <- stepAIC(model_with_interaction, direction = "both", family = binomial, k = log(nrow(bc_ref)))
summary(bic_selected_bin_model)


```
In the interaction model, the AIC selected model is status ~ race + marital_status + t_stage + n_stage + differentiate + estrogen_status + progesterone_status + reginol_node_positive + ln_tumor + sqrt_examined + age + race:marital_status + race:progesterone_status + race:ln_tumor.

The BIC selected model is status ~ race + differentiate + estrogen_status + progesterone_status + reginol_node_positive + ln_tumor + sqrt_examined + age.

### Model without Interaction Terms
```{r}
aic_selected_bin_model <- stepAIC(model_no_interaction, direction = "both")
summary(aic_selected_bin_model)

bic_selected_bin_model <- stepAIC(model_no_interaction, direction = "both", family = binomial, k = log(nrow(newbc)))
summary(bic_selected_bin_model)

```
In the model without interaction, the AIC selected model is status ~ n_stage + progesterone_status + differentiate + t_stage + age + race + estrogen_status + sqrt_examined + reginol_node_positive, having a AIC value of 2961.5.

The BIC selected model is status ~ progesterone_status + differentiate + t_stage + age + race + estrogen_status + sqrt_examined + reginol_node_positive with AIC value of 3003.7.




### Plots to check the model assumptions:

```{r plots check model assumption}


par(mfrow = c(3,4))
plot(aic_selected_bin_model)
plot(bic_selected_bin_model)


```








3.0 transformation edited
3.1 interaction transformation ??
3.2 partial test 
3.3 diagnostic boxcox
4. Step-wise: forward/ backward /AIC 
5. final model 
6. model assumption (check multicolinearity (VIF))
7. cross validation



