---
title: "project1"
date: "2023-12-18"
output: pdf_document
---

## Libraries

```{r library, message = FALSE}
library(tidyverse)
library(readr)
library(boot)
library(table1)
library(gridExtra)
library(MASS)
library(car)
library(dplyr)
library(leaps)
library(corrplot)
library(patchwork)
```

## Data Clean

```{r data load}
exam_score =
  read_csv("Project_1_data.csv") |>
  janitor::clean_names()
  
exam_score
summary(exam_score)
```



```{r data clean}
exam = exam_score |>
  mutate(
    gender = factor(gender, levels = c("female", "male")),
    ethnic_group = factor(ethnic_group, levels = c("group A", "group B", 
                                                       "group C", "group D", 
                                                       "group E")),
    parent_educ = factor(parent_educ, levels = c("some high school", "high school", 
                                                 "some college", "associate's degree", 
                                                 "bachelor's degree", "master's degree")),
    lunch_type = factor(lunch_type, levels = c("free/reduced", "standard")),
    test_prep = factor(test_prep, levels = c("none", "completed")),
    parent_marital_status = factor(parent_marital_status, levels = c("married", 
                                                     "divorced","single", "widowed")),
    practice_sport = factor(practice_sport, levels = c("never", "sometimes", "regularly")),
    is_first_child = factor(is_first_child, levels = c("yes", "no")),
    transport_means = factor(transport_means, levels = c("school_bus", "private")),
    wkly_study_hours = factor(wkly_study_hours, 
                                 levels = c("< 5", "> 10", "10-May"))
    )
exam
```
## Descriptive statistics for all variables

```{r Descriptive statistics}
summary(exam)
```




```{r data clean}
exam_cor = exam |>
  mutate(
    gender=case_when(
      gender == "female" ~ 0,
      gender == "male" ~ 1),
    ethnic_group=case_when(
      ethnic_group == "group A" ~ 1,
      ethnic_group == "group B" ~ 2,
      ethnic_group == "group C" ~ 3,
      ethnic_group == "group D" ~ 4,
      ethnic_group == "group E" ~ 5),
    parent_educ=case_when(
      parent_educ == "some high school" ~ 1,
      parent_educ == "high school" ~ 2,
      parent_educ == "some college" ~ 3,
      parent_educ == "associate's degree" ~ 4,
      parent_educ == "bachelor's degree" ~ 5,
      parent_educ == "master's degree" ~ 6),
    
    lunch_type=case_when(
      lunch_type == "free/reduced" ~ 1,
      lunch_type == "standard" ~ 2),
    test_prep=case_when(
      test_prep == "none" ~ 1,
      test_prep == "completed" ~ 2),
    parent_marital_status=case_when(
      parent_marital_status == "married" ~ 1,
      parent_marital_status == "divorced" ~ 2,
      parent_marital_status == "single" ~ 3,
      parent_marital_status == "widowed" ~ 4),
    practice_sport=case_when(
      practice_sport == "never" ~ 1,
      practice_sport == "sometimes" ~ 2,
      practice_sport == "regularly" ~ 3),
    is_first_child=case_when(
      is_first_child == "yes" ~ 1,
      is_first_child == "no" ~ 0),
    transport_means=case_when(
      transport_means == "school_bus" ~ 1,
      transport_means == "private" ~ 0),
    wkly_study_hours=case_when(
      wkly_study_hours == "< 5" ~ 1,
      wkly_study_hours == "> 10" ~ 2,
      wkly_study_hours == "10-May" ~ 3)

    ) |>
  na.omit()
```


## Covariance and Correlation

```{r correlation}
plot(exam_cor)

cor(exam_cor) |>
  knitr::kable(digits=4,caption="Correlation for all variables")

```

```{r plot for correlation}
corrplot(cor(exam_cor), method = "square", type = "upper", diag = FALSE)

```


## Exploratory visualisation



```{r}
plot_histograms_panels <- function(data, x_vars, y_vars, plots_per_panel = 9) {
  all_plots <- list()
  plot_count <- 1
  
  for (x in x_vars) {
    for (y in y_vars) {
      p <- data %>%
        ggplot(aes_string(x = x)) + 
        geom_bar(color = "blue", fill = alpha("blue", 0.5)) +
        theme_minimal() +
        labs(title = paste(y, "by", x),
             x = x,
             y = "Count")
      all_plots[[plot_count]] <- p
      plot_count <- plot_count + 1
    }
  }
  
  # Split the list of plots into multiple patchwork panels
  panel_list <- list()
  num_panels <- ceiling(length(all_plots) / plots_per_panel)
  for (i in 1:num_panels) {
    panel_start <- ((i - 1) * plots_per_panel) + 1
    panel_end <- min(i * plots_per_panel, length(all_plots))
    panel_plots <- all_plots[panel_start:panel_end]
    panel_list[[i]] <- do.call(patchwork::wrap_plots, c(panel_plots, ncol = 3))
  }
  
  return(panel_list)
}

# Example usage
x_vars <- c("gender", "ethnic_group", "parent_educ", "lunch_type","test_prep","parent_marital_status","practice_sport",
            "is_first_child","nr_siblings","transport_means", "wkly_study_hours")
y_vars <- c("reading_score", "writing_score", "math_score") # Replace with your y variables

panels <- plot_histograms_panels(exam, x_vars, y_vars, 9)

# To print the first panel, for example
print(panels[[1]])
print(panels[[2]])
print(panels[[3]])
print(panels[[4]])

```

# Statistical Testing

### T test for individual variables

For continuous variables and variables having only 2 categories.
```{r}
# Gender vs. 3 different Scores
t.test(math_score ~ gender, data = exam)
t.test(writing_score ~ gender, data = exam)
t.test(reading_score ~ gender, data = exam)

# Lunch-Type
t.test(math_score ~ lunch_type, data = exam)
t.test(writing_score ~ lunch_type, data = exam)
t.test(reading_score ~ lunch_type, data = exam)

# Test prep
t.test(math_score ~ test_prep, data = exam)
t.test(writing_score ~ test_prep, data = exam)
t.test(reading_score ~ test_prep, data = exam)

# is_first_child
t.test(math_score ~ is_first_child, data = exam)
t.test(writing_score ~ is_first_child, data = exam)
t.test(reading_score ~ is_first_child, data = exam)

# nr_siblings
cor.test(exam$math_score, exam$nr_siblings, use = "complete.obs")
cor.test(exam$reading_score, exam$nr_siblings, use = "complete.obs")
cor.test(exam$writing_score, exam$nr_siblings, use = "complete.obs")

# transport_means
t.test(math_score ~ transport_means, data = exam)
t.test(writing_score ~ transport_means, data = exam)
t.test(reading_score ~ transport_means, data = exam)



```



For variables with more than 2 categories

```{r}
perform_anova_tests <- function(data, x_vars, y_vars) {
  results <- list()
  result_count <- 1

  for (x in x_vars) {
    for (y in y_vars) {
      # Check if the variable has more than 2 levels
      if (length(unique(data[[x]])) > 2) {
        anova_result <- aov(reformulate(x, response = y), data = data)
        results[[result_count]] <- list(x_var = x, y_var = y, test_result = summary(anova_result))
        result_count <- result_count + 1
      }
    }
  }

  return(results)
}

# Example usage
x_vars <- c("ethnic_group", "parent_educ","parent_marital_status","practice_sport","wkly_study_hours") # Replace with your x variables
y_vars <- c("math_score", "reading_score", "writing_score") # Replace with your y variables

anova_results <- perform_anova_tests(exam, x_vars, y_vars)

# To print the results
for (result in anova_results) {
  print(paste("ANOVA for", result$x_var, "and", result$y_var))
  print(result$test_result)
}


```


From the result, we can see the p value is insignificant for all of those variables for Math score: is_first_child, nr_siblings, transportation_means, practice_sports

These variables are also insignificant for writing scores. 

For reading scores, only nb_siblings, transportation_means, and practice_sport are insignificant


## Model Building


```{r}
lm_math <- lm(math_score ~ gender + ethnic_group + parent_educ + lunch_type + test_prep + parent_marital_status + wkly_study_hours, data = exam)
summary(lm_math)

par(mfrow = c(2, 2))
plot(lm_math)
vif(lm_math)

```


```{r}
lm_writing <- lm(writing_score ~ gender + ethnic_group + parent_educ + lunch_type + test_prep + parent_marital_status + wkly_study_hours, data = exam)
summary(lm_writing)

par(mfrow = c(2, 2))
plot(lm_writing)
vif(lm_writing)

```


```{r}
lm_reading <- lm(math_score ~ gender + ethnic_group + parent_educ + lunch_type + test_prep + parent_marital_status + wkly_study_hours + is_first_child, data = exam)
summary(lm_reading)

par(mfrow = c(2, 2))
plot(lm_reading)

vif(lm_reading)

```



VIF and diagnostic seems good