---
title: "project2"
author: "Ze Li"
date: "2023-12-18"
output: pdf_document
---

## Libraries

```{r library, message = FALSE}
library(tidyverse)
library(readr)
library(MASS)
library(car)
library(dplyr)
library(leaps)
library(survival)
library(survminer)
```

## Data Clean

```{r data load}
breastcancer_data =
  read_csv("Project_2_data.csv") |>
  janitor::clean_names()
```

```{r data clean}
bc = breastcancer_data |>
  mutate(
    race = factor(race, levels = c("White", "Black", "Other")),
    marital_status = factor(marital_status, levels = c("Married", "Divorced", 
                                                       "Single", "Widowed", 
                                                       "Separated")),
    t_stage = factor(t_stage, levels = c("T1", "T2", "T3", "T4")),
    n_stage = factor(n_stage, levels = c("N1", "N2", "N3")),
    x6th_stage = factor(x6th_stage, levels = c("IIA", "IIIA", "IIIC", "IIB", "IIIB")),
    differentiate = factor(differentiate, levels = c("Poorly differentiated", 
                                                     "Moderately differentiated", 
                                                     "Well differentiated", 
                                                     "Undifferentiated")),
    grade = factor(grade, levels = c("1", "2", "3", "anaplastic; Grade IV")),
    a_stage = factor(a_stage, levels = c("Distant", "Regional")),
    estrogen_status = factor(estrogen_status, levels = c("Negative", "Positive")),
    progesterone_status = factor(progesterone_status, 
                                 levels = c("Negative", "Positive")),
    status = factor(status, levels = c("Dead", "Alive"))
  )
```

## Descriptive statistics for all variables

```{r Descriptive statistics}
summary(bc)
```

### Kaplan-Meier survival curves

```{r km plot}
bc$survival_months <- as.numeric(bc$survival_months)
bc$status <- as.numeric(bc$status)
surv_obj <- Surv(time = bc$survival_months, event = bc$status)

km_fit <- survfit(surv_obj ~ race, data = bc)
plot(km_fit, col = 1:3, xlab = "Time", ylab = "Survival Probability")
legend("bottomleft", legend = levels(bc$race), col = 1:3, lty = 1)

ggsurvplot(km_fit, data = bc, pval = TRUE, conf.int = TRUE,
           xlab = "Time", ylab = "Survival Probability",
           risk.table = TRUE, # Add risk table
           risk.table.height = 0.3, # Adjust the height of the risk table
           risk.table.fontsize = 2, # Adjust the font size in the risk table
           conf.int.size = 0.5, # Adjust the size of the confidence interval lines
           palette = c("#00AFBB", "#E7B800", "#FC4E07"))
```

The Kaplan-Meier plot displays survival probabilities for three racial groups: White, Black, and Other, over time. Survival curves are closely aligned, suggesting minimal differences in survival outcomes across these groups, which is statistically supported by a p-value of 0.73, indicating no significant difference. The number-at-risk table below the plot shows a decreasing count over time, reflecting those still under observation at specific time points, with the White group starting with the highest count. 

Therefore, the consistency of the curves along with the non-significant p-value suggests race, in this sample, does not have a differential impact on survival probability.

## Fit a Cox Proportional Hazards Model

```{r cox model}
bc$survival_months <- as.numeric(bc$survival_months)
bc$status <- as.numeric(bc$status)
surv_obj <- Surv(time = bc$survival_months, event = bc$status)

# Fit the Cox model
cox_model <- coxph(surv_obj ~ age + race + marital_status + t_stage + n_stage + x6th_stage + 
                     differentiate + grade + a_stage + tumor_size + estrogen_status + 
                     progesterone_status + regional_node_examined + reginol_node_positive, data = bc)
summary(cox_model)
```

Concordance= 0.538  (se = 0.006 )

### Check Proportional Hazards Assumption

```{r Schoenfeld residuals}
zph <- cox.zph(cox_model)
plot(zph)
```

A non-random pattern or a significant global test (p-value) can indicate violations of the assumption.

### Variable Selection 

```{r stepwise}
stepwise_model <- stepAIC(cox_model, direction = "both", trace = FALSE)
stepwise_model
summary(stepwise_model)
```

After stepwise, it kept two variables differentiate and progesterone status.

Concordance= 0.525  (se = 0.005 )

### Model Diagnostics

```{r diagnostics}
# Identify Influential Observations
martingale_res <- residuals(cox_model, type = "martingale")
plot(martingale_res)

martingale_res2 <- residuals(stepwise_model, type = "martingale")
plot(martingale_res2)
```

It seems both of plots have few observations with large negative Martingale residuals, which could be potential outliers or influential observations.  

### Model Validation

```{r stepwise validation}
y <- bc[["status"]] # Target column
trainIndex <- caret::createDataPartition(y, p = 0.9, list = FALSE)

# Create training and testing sets
train=bc[trainIndex,]
test=bc[-trainIndex,]

# train
surv_obj <- Surv(time = train$survival_months, event = train$status)
fit <- coxph(surv_obj ~ differentiate + progesterone_status, data = train)
summary(fit)

# Test data
predictions <- predict(stepwise_model, newdata = test, type = "risk")
perf_metric <- survival::survConcordance(Surv(survival_months, status) ~ predictions, 
                                         data = test)$concordance
perf_metric
```

concordant 
 0.5336526 

```{r full model validation}
y <- bc[["status"]] # Target column
trainIndex <- caret::createDataPartition(y, p = 0.9, list = FALSE)

# Create training and testing sets
train=bc[trainIndex,]
test=bc[-trainIndex,]

# train
surv_obj <- Surv(time = train$survival_months, event = train$status)
fit <- coxph(surv_obj ~ age + race + marital_status + t_stage + n_stage + x6th_stage + 
                     differentiate + grade + a_stage + tumor_size + estrogen_status + 
                     progesterone_status + regional_node_examined + reginol_node_positive, 
             data = train)
summary(fit)

# Test data
predictions <- predict(stepwise_model, newdata = test, type = "risk")
perf_metric <- survival::survConcordance(Surv(survival_months, status) ~ predictions, 
                                         data = test)$concordance
perf_metric
```

concordant 
 0.5430644 